Implicit assumptions and their impact on photometric redshift PDF performance in the context of LSST

### 2018-12-03 paper/main.tex
edits based on read-through of paper

In order to maximize scientific returns of current and upcoming galaxy surveys, the photometric redshift (photo-$z$) posterior distributions produced by redshift estimation codes must be accurate probability distribution functions (PDFs).  However, the posteriors resulting from a number of current techniques are not, in general, consistent with each other, affected by implicit assumptions made by each code, and an optimal method for obtaining an accurate PDF estimate remains unclear.   We present the results of an initial study of the Large Synoptic Survey Telescope Dark Energy Science Collaboration (\textsc{LSST-DESC}) evaluating twelve photo-$z$ algorithms using complete and representative training data and evaluate multiple metrics to test how accurately the posteriors represent probability distributions.  We observe several trends, including systematic biases and an overall over/under-prediction in the broadness of the PDFs in many of the codes which may be symptomatic of implementation problems or problems in underlying algorithm design.  A careful accounting of all systematics discovered will be necessary for the codes employed in upcoming analyses in order to achieve unbiased cosmological measurements.

>>>>

In order to maximize scientific returns of current and upcoming galaxy surveys, the photometric redshift (photo-$z$) posterior distributions produced by redshift estimation codes must be accurate probability distribution functions (PDFs).  However, the posteriors resulting from a number of current techniques are not, in general, consistent with each other, affected by implicit assumptions made by each code, and an optimal method for obtaining an accurate PDF estimate remains unclear.   We present the results of an initial study of the Large Synoptic Survey Telescope Dark Energy Science Collaboration (\textsc{LSST-DESC}) testing twelve photo-$z$ algorithms using complete and representative training data and evaluate multiple metrics to test how accurately the posteriors represent probability distributions.  We observe several trends, including systematic biases and an overall over/under-prediction in the broadness of the PDFs in many of the codes which may be symptomatic of implementation problems or problems in underlying algorithm design.  A careful accounting of all systematics discovered will be necessary for the codes employed in upcoming analyses in order to achieve unbiased cosmological measurements.

### 2018-12-03 paper/tex/discussion.tex
edits based on read-through of paper

\section{Discussion and Future Work}
\label{sec:discussion}
%(Sam Schmidt, Ofer Lahav, Jeff Newman, Alex Malz, Eve Kovacs, Tony Tyson, Tina Peters)
In this paper we presented results evaluating the computation of individual galaxy photometric redshift PDFs for twelve photo-$z$ codes.  As discussed in Section~\ref{sec:metrics} each $p(z)$ should accurately reflect the relative likelihood as a function of redshift for each galaxy in an informative way; that is, the estimates should provide useful information per individual galaxy, not just the ensemble.  All codes were provided a set of representative training data and tested on an idealized set of model galaxies with high signal-to-noise and photometry with no confounding effects due to blending, instrumental effects, the night sky, or other complications included.  The goal was not to determine a ``best'' photo-$z$ code: in many ways, this was a baseline test of a ``best case scenario'' to predict the expected photo-$z$ performance if a stage IV dark energy survey was to obtain complete training samples and perfectly calibrated their multi-band photometry.  Given these idealized conditions, any deficiencies observed in a photo-$z$ code's performance should be a cause for concern, and may be evidence of a problem with either/both of the specific code implementation or the underlying algorithm.  In order to meet the stringent LSST goals for photo-$z$ performance, identifying and correcting such problems is an important first step before tackling more realistic data in future challenges, and codes that do not perform well may not be worth pursuing in future challenges.  Many of the codes tested performed well; however, several did not meet the stringent goals that have been laid out for LSST photometric redshift performance for individual galaxies, as laid out in the LSST SRD (See Section~\ref{sec:intro}.  This is a cause for concern, given the idealized conditions, and the individual code responses will be studied in detail moving forward.  If methods can not reach the goals on idealized data, then they will almost surely not meet those same goals when the more complex problems that we expect to arise from real LSST data are included.  The results presented in this paper enable an evaluation of which algorithms are the most promising moving forward, and potentially point to implementation choices or mistakes which could be improved or corrected in others.

@@

Another important factor to keep in mind when examining the results presented in this paper is the fact that they are at some level dependent on the metrics that we aim to optimize: in this case code participants were asked to submit their optimal measures of an accurate $p(z)$, so participants used the training/validation data to optimize their codes accordingly.  Had we, instead, asked for an optimal $\hat{N(z)}$ the resulting metrics would be different for most, if not all, of the codes, as they would optimize toward a different goal.  Specific metric choice can affect which codes are among the ``best'' codes.  As stated earlier, there are cosmological science cases that require either individual galaxy photo-$z$ measures, or ensemble $\hat{N}(z)$ measures.  We must be aware of that the optimal method for one is not necessarily optimal for the other, and in fact several photo-$z$ algorithms may be necessary in the final cosmological analysis in order to satisfy the requirements of all science use cases.  The example of the simple \trainz\ estimator described in Section~\ref{sec:caution} shows a simple model with a $p(z)$ that is unrealistic for individual objects can still score very well on many of our metrics.  It is important to look at {\it all} metrics, and keep in mind what information each metric conveys.
%\red{mention 'null' and how 'dumb' algorithms can score well on metrics, need to keep all metrics in mind}
%\aim{[reiterate the meaning of $p(z)$ and the goal -- emphasize finding ``best'' code in terms of impact of assumptions underlying the method \textit{or} establishing methodology for the realistic case of biased prior information and other science goals]}
%\red{[the challenges faced in this project]}
%\red{mention z$<$2, missing some important degeneracies, new sim will go to higher z}
%\red{no quality cuts included, could identify outliers, but could also induce biases}
%\red{[how optimization defined has impact, here we optimized p(z), not n(z), could be science-case specific if stringent requirements are needed, though that may be a computational/storage challenge if too many use cases needed.]}
%\red{some discussion of output parameterization, limitations of 200 point fixed grid, in future switch to something like quantiles (cite qp paper)}
We re-emphasize that the dataset tested was quite idealized, and discuss enhancements that will be added in future simulations to test photo-$z$ codes on increasingly realistic conditions in the following section.

>>>>

\section{Discussion and Future Work}
\label{sec:discussion}
%(Sam Schmidt, Ofer Lahav, Jeff Newman, Alex Malz, Eve Kovacs, Tony Tyson, Tina Peters)
In this paper we presented results evaluating the computation of individual galaxy photometric redshift PDFs for twelve photo-$z$ codes.  As discussed in Section~\ref{sec:metrics} each $p(z)$ should accurately reflect the relative likelihood as a function of redshift for each galaxy in an informative way; that is, the estimates should provide useful information per individual galaxy, not just the ensemble.  All codes were provided a set of representative training data and tested on an idealized set of model galaxies with high signal-to-noise and photometry with no confounding effects due to blending, instrumental effects, the night sky, or other complications included.  %The goal was not to determine a ``best'' photo-$z$ code:i
In many ways, this was a baseline test of a ``best case scenario'' to predict the expected photo-$z$ performance if a stage IV dark energy survey was to obtain complete training samples and perfectly calibrated their multi-band photometry.  Given these idealized conditions, any deficiencies observed in a photo-$z$ code's performance should be a cause for concern, and may be evidence of a problem with either/both of the specific code implementation or the underlying algorithm.
Many of the codes tested performed well; however, several did not meet the stringent goals that have been laid out for LSST photometric redshift performance for individual galaxies, as laid out in the LSST SRD (See Section~\ref{sec:intro}).%  This is a cause for concern, and the individual code responses will be studied in detail moving forward.
If methods can not reach the goals on idealized data, then they will almost surely not meet those same goals when the more complex problems that we expect to arise from real LSST data are included.  The results presented in this paper enable an evaluation of which algorithms are the most promising moving forward, and potentially point to implementation choices or mistakes which could be improved or corrected in others.  Codes that do not perform well may not be worth pursuing in future challenges.

@@

Another important factor to keep in mind when examining the results presented in this paper is the fact that they are at some level dependent on the metrics that we aim to optimize: in this case code participants were asked to submit their optimal measures of an accurate $p(z)$, so participants used the training/validation data to optimize their codes accordingly.  Had we, instead, asked for an optimal $\hat{N(z)}$ the resulting metrics would be different for most, if not all, of the codes, as they would optimize toward a different goal.  Specific metric choice can affect which codes are among the ``best'' codes.  As stated earlier, there are cosmological science cases that require either individual galaxy photo-$z$ measures, or ensemble $\hat{N}(z)$ measures.  We must be aware of that the optimal method for one is not necessarily optimal for the other, and in fact several photo-$z$ algorithms may be necessary in the final cosmological analysis in order to satisfy the requirements of all science use cases.

The example of the simple \trainz\ estimator described in Section~\ref{sec:caution} shows a simple model with a $p(z)$ that is unrealistic for individual objects can still score very well on many of our metrics.  It is important to look at {\it all} metrics, and keep in mind what information each metric conveys.
%\red{mention 'null' and how 'dumb' algorithms can score well on metrics, need to keep all metrics in mind}
%\aim{[reiterate the meaning of $p(z)$ and the goal -- emphasize finding ``best'' code in terms of impact of assumptions underlying the method \textit{or} establishing methodology for the realistic case of biased prior information and other science goals]}
%\red{[the challenges faced in this project]}
%\red{mention z$<$2, missing some important degeneracies, new sim will go to higher z}
%\red{no quality cuts included, could identify outliers, but could also induce biases}
%\red{[how optimization defined has impact, here we optimized p(z), not n(z), could be science-case specific if stringent requirements are needed, though that may be a computational/storage challenge if too many use cases needed.]}
%\red{some discussion of output parameterization, limitations of 200 point fixed grid, in future switch to something like quantiles (cite qp paper)}
We re-emphasize that the dataset tested was quite idealized, and discuss enhancements that will be added in future simulations to test photo-$z$ codes on increasingly realistic conditions in the following section.

### 2018-12-03 paper/tex/introduction.tex
edits based on read-through of paper

The unprecedented sample size of LSST galaxies, expected to number several billion for the main cosmological sample, necessitates stringent constraints on photo-$z$ accuracy if systematic errors are not to dominate the statistical errors.
The LSST Science Requirements Document (SRD)\footnote{available at \url{https://docushare.lsstcorp.org/docushare/dsweb/Get/LPM-17}} lists the individual galaxy photometric redshift goals for a magnitude limited sample with $i<25$ as: root-mean-square error with a goal of $\sigma_z<0.02(1+z)$; $3\sigma$ ``catastrophic outlier'' rate below $10\%$; bias below $0.003$\footnote{
Note that at the time the SRD was written, these goals were stated in terms of a photo-$z$ point estimate for each galaxy, as was standard in many previous studies, while in this paper we emphasize the importance of using a full photo-$z$ PDF.}.  The LSST Dark Energy Science Collaboration (LSST-DESC) developed a separate Science Requirements Document \citep{Mandelbaum:2018}, which forecasts the constraining power of five cosmological probes using somewhat conservative assumptions to define requirements on systematic errors for several measurements.  These include even more stringent requirements on photometric redshift performance than those included in the LSST SRD, though most of the initial LSST-DESC requirements are defined in terms of tomographic bin populations rather than on individual object redshifts.  The tremendous size of LSST's galaxy catalogue will be enabled by its exceptional depth, pushing to fainter magnitudes and deeper imaging and including galaxies of lower luminosity and higher redshift than ever before.
The inclusion of these populations introduce major physical degeneracies, for example the Lyman break/Balmer break degeneracy, that were not present in the populations covered in shallower large area surveys like the Sloan Digital Sky Survey \citep[SDSS,][]{York:00} and the Two Micron All Sky Survey \citep[2MASS,][]{Skrutskie:06}.  These issues are not unique to LSST, and are present in Stage III Dark Energy surveys; however,
in order to meet the demanding error budgets of Stage IV projects such as LSST and LSST-DESC it will be necessary to fully characterize those degeneracies wherein multiple redshift solutions have comparable likelihood to per cent level accuracy.

@@

There are numerous techniques for deriving photo-$z$ PDFs from photometry, yet no one method has yet been established as clearly superior.
Quantitative comparisons of photo-$z$ methods have been made before.
The Photo-$z$ Accuracy And Testing \citep[PHAT,][]{Hildebrandt:10} effort focused on point estimates derived from many photometric bands.  \citet{Rau:2015} introduced a new method for improving redshift PDFs using an ordinal classification algorithm.
DES compared several codes for point estimates and a subset with $p(z)$ information \citep{Sanchez:14}.  A follup up paper examined summary statistics of photo-$z$ interim posteriors for tomographically binned galaxy subsamples \citep{Bonnett:16}.

This paper is distinguished by its focus on metrics of photo-$z$ interim posteriors themselves and consideration of both classic and state-of-the-art photo-$z$ algorithms, comparing the performance of several of the most widely employed codes as well as some that have been developed only recently on the basis of metrics appropriate for a probabilistic data product.
The results presented here are
%part of the Data Challenge 1 (DC1)
a major focus of the Photometric Redshift working group of the LSST-DESC.
This work is laid out in the Science Roadmap (SRM)\footnote{Available at: \url{http://lsst-desc.org/sites/default/files/DESC_SRM_V1_1.pdf}} as one of the critical activities to be completed in preparation for dark energy science analysis on the first year LSST data.
%This is the first of multiple papers by the working group, which will grow in sophistication.
In this initial paper we focus on evaluating the performance of photometric redshift codes and PDF-based performance metrics in the presence of complete and representative training sets.  Specific implementation choices in each code will influence the resultant posterior distributions, for example choice of prior parameterization in template-based codes, the bandwidth size chosen for machine learning based codes, or even the output format chosen for storing the PDF.  We have attempted to minimize the impact of many of these factors when comparing codes, for example by using the same template set for all template-based codes, and using a training set that is drawn from the same underlying population as the test sample, to create a controlled environment in which to compare the photo-$z$ PDFs derived from each method.  We explore a number of performance metrics in this paper that test whether the posterior estimates are actual PDFs.  
Comparing the relative performance of the codes enables us to evaluate whether each code is using information in an optimal way, and may reveal enhancements in some codes and deficiencies in others, either in the fundamental algorithm, or in specific implementation.  Identifying and fixing failure modes within codes may aid us in reaching the stringent photo-$z$ performance goals set out for LSST.  We note that these initial tests are a necessary requirement for photo-$z$ codes that will be used in cosmological analyses; however, meeting these requirements is only the first stage in the process, and can be thought of as an initial test under near perfect conditions to test for problems before further complexities are added in future analyses.

>>>>

The unprecedented sample size of LSST, expected to number several billion galaxies for the main cosmological sample, necessitates stringent constraints on photo-$z$ accuracy if systematic errors are not to dominate the statistical errors.
The LSST Science Requirements Document (SRD)\footnote{available at \url{https://docushare.lsstcorp.org/docushare/dsweb/Get/LPM-17}} lists the individual galaxy photometric redshift goals for a magnitude limited sample with $i<25$ as: root-mean-square error with a goal of $\sigma_z<0.02(1+z)$; $3\sigma$ ``catastrophic outlier'' rate below $10\%$; bias below $0.003$\footnote{
Note that at the time the SRD was written, these goals were stated in terms of a photo-$z$ point estimate for each galaxy, as was standard in many previous studies, while in this paper we emphasize the importance of using a full photo-$z$ PDF.}.  The LSST Dark Energy Science Collaboration (LSST-DESC) developed a separate Science Requirements Document \citep{Mandelbaum:2018}, which forecasts the constraining power of five cosmological probes using somewhat conservative assumptions to define requirements on systematic errors for several measurements.  These include even more stringent requirements on photometric redshift performance than those included in the LSST SRD, though most of the current LSST-DESC requirements are defined in terms of tomographic bin populations rather than on individual object redshifts.  The tremendous size of LSST's galaxy catalogue will be enabled by its exceptional depth, pushing to fainter magnitudes and deeper imaging and including galaxies of lower luminosity and higher redshift than ever before.
The inclusion of these populations introduce major physical degeneracies, for example the Lyman break/Balmer break degeneracy, that were not present in the populations covered in shallower large area surveys like the Sloan Digital Sky Survey \citep[SDSS,][]{York:00} and the Two Micron All Sky Survey \citep[2MASS,][]{Skrutskie:06}.  These issues are not unique to LSST, and are present in Stage III Dark Energy surveys; however,
in order to meet the demanding error budgets of Stage IV projects such as LSST and LSST-DESC it will be necessary to fully characterize those degeneracies wherein multiple redshift solutions have comparable likelihood to per cent level accuracy.

@@

There are numerous techniques for deriving photo-$z$ PDFs from photometry, yet no one method has yet been established as clearly superior.
Quantitative comparisons of photo-$z$ methods have been made before.
The Photo-$z$ Accuracy And Testing \citep[PHAT,][]{Hildebrandt:10} effort focused on point estimates derived from many photometric bands.  \citet{Rau:2015} introduced a new method for improving redshift PDFs using an ordinal classification algorithm.
DES compared several codes for point estimates and a subset with $p(z)$ information \citep{Sanchez:14}.  A follow up up paper examined summary statistics of photo-$z$ interim posteriors for tomographically binned galaxy subsamples \citep{Bonnett:16}.

This paper is distinguished by its focus on metrics of photo-$z$ interim posteriors themselves and consideration of both classic and state-of-the-art photo-$z$ algorithms, comparing the performance of several of the most widely employed codes, as well as some that have been developed only recently, on the basis of metrics appropriate for a probabilistic data product.
The results presented here are
%part of the Data Challenge 1 (DC1)
a major focus of the Photometric Redshift working group of the LSST-DESC.
This work is laid out in the Science Roadmap (SRM)\footnote{Available at: \url{http://lsst-desc.org/sites/default/files/DESC_SRM_V1_1.pdf}} as one of the critical activities to be completed in preparation for dark energy science analysis on the first year LSST data.
%This is the first of multiple papers by the working group, which will grow in sophistication.
In this initial paper we focus on evaluating the performance of photometric redshift codes and PDF-based performance metrics in the presence of complete and representative training sets.  Specific implementation choices in each code will influence the resultant posterior distributions, for example choice of prior parameterization in template-based codes, the bandwidth size chosen for machine learning based codes, or even the output format chosen for storing the PDF.  We have attempted to minimize the impact of many of these factors when comparing codes, for example by using the same template set for all template-based codes, and using a training set that is drawn from the same underlying population as the test sample, to create a controlled environment in which to compare the photo-$z$ PDFs derived from each method.  We explore a number of performance metrics in this paper that test whether the posterior estimates are actual PDFs.  
Comparing the relative performance of the codes enables us to evaluate whether each code is using information in an optimal way, and may reveal enhancements in some codes and deficiencies in others, either in the fundamental algorithm, or in specific implementation.  Identifying and fixing failure modes within codes may aid us in reaching the stringent photo-$z$ performance goals set out for LSST.  We note that these initial tests are a necessary requirement for photo-$z$ codes that will be used in cosmological analyses; however, meeting these requirements is only the first stage in the process, and can be thought of as an initial test under near perfect conditions to test for problems before further complexities are added in future analyses.
